---
title: "Exploratory Analysis"
author: "AJ Catalano"
date: "11/9/2021"
output: github_document
---

```{r}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = .6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.color = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Exploratory analysis includes:

* visualization 
* checks for data completion and reliability
* initial hypothesis evaluation
* hypothesis formation

**Want numerical summaries of data, especially within groups and subgroups**

```{r}
weather_df =  
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) %>%
  select(name, id, everything())
```

## `group_by`

```{r}
# `ungroup` does what it sounds like it does

weather_df %>% 
  group_by(name, month)
```

## Counting

```{r}
# counting month observations; n() gives current group size

weather_df %>% 
  group_by(month) %>% 
  summarize(n_obs = n())

weather_df %>% 
  group_by(name, month) %>% 
  summarize(n_obs = n())
```

`count()`

```{r}
weather_df %>% 
  count(name, month, name = "n_obs")
```

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    n_obs = n(),
    n_days = n_distinct(date)
  )
```

## 2x2 tables

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE     ~ ""
    )
  ) %>% 
  group_by(name, cold) %>% 
  summarize(count = n())
```

```{r}
weather_df %>% 
  filter(name != "Waikiki_HA") %>% 
  mutate(
    cold = case_when(
      tmax < 5 ~ "cold",
      tmax >= 5 ~ "not_cold",
      TRUE     ~ ""
    )
  ) %>% 
  janitor::tabyl(name, cold)
```

## General summaries

```{r}
weather_df %>% 
  group_by(month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  )
```

**The output of the above code is a dataframe...and can be piped into ggplot!**

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    mean_prcp = mean(prcp, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

Summarizing many columns using `across`:

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(across(prcp:tmin, mean))
```

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(mean_tmax = mean(tmax)) %>% 
  pivot_wider(names_from = name, values_from = mean_tmax) %>% 
  knitr::kable(digits = 1)

# knitr::kable makes tables look better when knit
```

## `group_by` and `mutate`

```{r}
# dataframe will be same size as weather_df, but will have a new mean_tmax and centered_tmax column
weather_df %>% 
  group_by(name) %>% 
  mutate(
    mean_tmax = mean(tmax, na.rm = TRUE),
    centered_tmax = tmax - mean_tmax
  ) %>% 
  ggplot(aes(x = date, y = centered_tmax, color = name)) +
  geom_point()
```

## Window functions

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    temp_rank = min_rank(tmax)
  ) %>% 
  filter(temp_rank == 1) %>%
  view()

# ranking in descending order

weather_df %>% 
  group_by(name, month) %>% 
  mutate(
    temp_rank = min_rank(desc(tmax))
  ) %>% 
  view()
```

## Lag

```{r}
weather_df %>% 
  group_by(name) %>% 
  mutate(
    lag_temp = lag(tmax, 5),
    temp_change = tmax - lag(tmax)
  ) %>% 
  summarize(
    temp_change_max = max(temp_change, na.rm = TRUE),
    temp_change_sd = sd(temp_change, na.rm = TRUE)
  )
```

