---
title: "Data Visualization"
author: "AJ Catalano"
date: "11/7/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

Exploratory analysis may include:

* visualization
* checks for data completeness and reliablity
* quantification of centrality and variabilty
* initial evaluation of hypotheses
* hypothesis generation

**Basic graph components:**

* data
* aesthetic mapping - what variables map to what features of the graph (e.g. axes, point size, colors)
* geom - visual representaion of data (type of plot)

**Advanced graph components:**

* facets - multiple panels
* scales
* statistics

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplots

```{r}
ggplot(
  weather_df, aes(x = tmin, y = tmax)) + 
    geom_point()

# different approach to same plot:

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Saving and editing plot object:

```{r}
# first:

weather_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

# then:

weather_plot + geom_point()
```

