---
title: "Data Visualization"
author: "AJ Catalano"
date: "11/7/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
```

Exploratory analysis may include:

* visualization
* checks for data completeness and reliablity
* quantification of centrality and variabilty
* initial evaluation of hypotheses
* hypothesis generation

**Basic graph components:**

* data
* aesthetic mapping - what variables map to what features of the graph (e.g. axes, point size, colors)
* geom - visual representaion of data (type of plot)

**Advanced graph components:**

* facets - multiple panels
* scales
* statistics

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplots

```{r}
ggplot(
  weather_df, aes(x = tmin, y = tmax)) + 
    geom_point()

# different approach to same plot:

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```

Saving and editing plot object:

```{r}
# first:

weather_plot =
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

# then:

weather_plot + geom_point()
```

## Advanced scatterplots

```{r}
# adding color and smooth trend lines

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE)
```

What about `aes` placement?

```{r}
# note that placing the color aes within `geom_point` places a single smooth trend line through the scatterplot. R is reading the color aes within the whole scatterplot

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name)) + 
  geom_smooth(se = FALSE)
```

## Faceting

```{r}
# `facet_grid` first argument = row definition second argument = column definition

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.2, size = 0.3) + 
  geom_smooth(se = FALSE, size = 2) +
  facet_grid(. ~ name)

# below alpha = tmin specifies a color gradient mapped to variable tmin

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, alpha = tmin, color = name)) +
  geom_point() + 
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .)
```

Combining elements:

```{r}
# size of points is mapped to precipitation amount

weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  theme_classic()
```

How many geoms must exist?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

Neat geoms exist!

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_hex()

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_density_2d() +
  geom_point(alpha = 0.3)
```

## Univariate plots

Histograms

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Adding color and stuff (can be weird):

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")

# using facets can make multiple levels easier to see

weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

A new geometry (density = smoothed histogram, loses outliers - adjust with `adjust`):

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.3, adjust = 0.5)
```

## Box plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

## Trendy plots

Violin plots:

```{r}
# turning your head sideways lets you see densities (makes skewed data more obvious than box plots)

weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) + 
  geom_violin(alpha = 0.5) +
  stat_summary(fun = median)
```

Ridge plots (good for many categories e.g. age distribution by US state):

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

## Saving and embeding plots

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5)

ggsave("./results/weather_plot.pdf", weather_plot, width = 8, height = 5)
```

Embedding:

```{r fig.width=6, fig.height=3}
weather_plot
```

```{r fig.width=12, fig.height=6}
weather_plot
```

